---
title: "Topic modeling and visualization"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Topic modeling and visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  message = FALSE,
  # error = FALSE,
  warning = FALSE,
  fig.align = "center"
)
options(width = 80)
```

```{r}
library(CellFunMap)
library(magrittr)
```

First, please make sure that you have previously performed the pre-processing and GSEA steps, see [Pre-processing](Pre-processing.html).

```{r, echo = FALSE}
SeuratObj <- readRDS(file = "../vignette_data/pbmc3k_SeuratObj.rds")
```

## Topic modeling

Latent Dirichlet Allocation (LDA) is a Bayesian mixture model, initially developed for the analysis of text documents. **CellFunMap** provides a method to fit an LDA model to the clusters-by-pathways matrix derived from GSEA, where each cluster is considered as a document, pathway enrichment score(log-transformed) as a token in the document, and each pathway is considered as a word. CellFunMap infers the topic probability distribution by decomposing the clusters-by-pathways matrix into clusters-by-topics (`θ`) and topics-by-pathways (`ϕ`) matrices, that is, (1) the probability distributions over the topics for each cluster in the dataset (`θ`) and (2) the probability distributions over the pathways for each topic (`ϕ`). These distributions indicate, respectively, how important a regulatory topic is for a cluster (`θ`), and how important pathways are for the regulatory topic (`ϕ`). We assume that each cluster can be represented as a mixture of latent topics, which reveal cellular programs shared across cell types or exclusive to a particular cell type.

Actually, CellFunMap provides two metrics, **log-likelihood** and **perplexity**, to determine the optimal number of topics. But we recommend a customized number of topics such as the number of cell types in scRNA-seq data, in order to improve biological interpretability. 

```{r, eval = FALSE}
k <- length(unique(Seurat::Idents(SeuratObj)))
# if k=NULL, the optimal number of topics will be calculated automatically according to log-likelihood and perplexity.
SeuratObj <- runLDA(SeuratObj, by = "GO", k = k, method = "VEM", SEED = 1234, plot = F)
```

```{r, eval = FALSE}
# save object for later use.
save(SeuratObj, file = "SeuratObj.RData")
```

## Visualization of topic modelling results

CellFunMap provides a few methods to visualize topic modelling results.

```{r}
ldaOut <- SeuratObj@misc$ldaOut
betaDF <- tidytext::tidy(ldaOut, matrix = "beta")
pws <- ID2Description(SeuratObj, by = "GO")
betaDF %<>% dplyr::mutate(descrip=unname(pws[term]))
gammaDF <- tidytext::tidy(ldaOut, matrix = "gamma")
```

```{r, fig.height=6, fig.width=7}
# Show topic distribution on UMAP/TSNE cell map
topicProb(SeuratObj, reduction="umap", topic=3, pointSize=0.1)
```

```{r, fig.height=11, fig.width=8}
# echart4r network showing top terms of topics with tooltips showing description of pathway ID
topicNW3(betaDF, topn=10, pws = pws)
```

```{r, fig.height=7, fig.width=7}
# Heatmap showing probability between topics and clusters
cluster_topic_hmp(ldaOut)
```

```{r, fig.height=10, fig.width=10}
# Sankey diagram showing the best assigned topic of each cluster
plot_sankey(gammaDF, topn=1)
```

```{r, fig.height=7, fig.width=7}
# word cloud of top terms of topic
wordcloud_topic(betaDF, pws, topic=3, topn=20)
```

```{r, fig.height=7, fig.width=7}
# 3D word cloud of top terms of topic
wordcloud_topic_3D(betaDF, pws, topic=3, topn=20)
```

```{r, fig.height=7, fig.width=7}
# Heatmap showing cosine similarity between topics
cosineheatmap(ldaOut)
```

```{r}
# Heatmap showing cosine similarity between clusters
cosinehmp_cluster(ldaOut)
```

```{r, fig.height=7, fig.width=7}
# Network showing cosine similarity between clusters
cosine_network_cluster(ldaOut, layout="fr", cos_sim_thresh=0.02, SEED=123, radius=0.12, width_range=c(0.1, 0.8))
```

```{r, fig.height=10, fig.width=10}
# Network of cosine similarity between terms
cosine_network_term(SeuratObj, cosine_cal_by = "Topic modeling", pie_by = "Topic modeling", GSEA_by = "GO",
   topn = 10, layout = "fr", cos_sim_thresh = 0.8, radius = 0.1, text_size = 2)
```

```{r, fig.height=6, fig.width=7}
# UMAP on cluster-topic probability matrix
umap_cluster(ldaOut)
```

```{r, fig.height=7, fig.width=7}
# Show top terms of each topic
Topterms_Topic(betaDF, Topic = 3, topn = 10, text.size = 3)
```

```{r, fig.height=10, fig.width=10}
# barplot to show top terms of every topic
hist_topic_term(betaDF, topics=c(1,2,3,4), topn=20, axis.text.y.size=3)
```

```{r, fig.height=10, fig.width=10}
# barplot to show probability of assigned topics in clusters
hist_cluster_topic(gammaDF, clusters=head(Seurat::Idents(SeuratObj)))
```


