---
title: "Transfer annotation from reference to query data"
output: html_document
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  message = FALSE,
  # error = FALSE,
  warning = FALSE,
  fig.align = "center"
)
options(width = 80)
```

**CellFunMap** provides methods for rapid, automated cell-type annotation across datasets without dependence on marker genes, reducing labor intensive and time-consuming work of manual markers collection and manual cell-type annotation.

## Predict cell type of query data based on reference data

```{r}
library(CellFunMap)
```

Here, we take `pbmc3k.final` as the reference data:

```{r}
data('pbmc3k.final', package = "pbmc3k.SeuratData")
unique(Seurat::Idents(pbmc3k.final))
```

To make things easier, we re-analyze and re-cluster `pbmc3k.final` as the query data, each cell owns a new cluster identity.

```{r}
SeuratObj <- readData(data = pbmc3k.final, type = 'Seurat', species = "Homo sapiens")
SeuratObj <- CalMTpercent(SeuratObj, by = "use_internal_data")
SeuratObj <- QCfun(SeuratObj, plot = F)
SeuratObj <- RunSeurat(SeuratObj, nPCs = 10, resolution = 1, plot = FALSE)
SeuratObj <- RunGSEA(SeuratObj, by = 'GO')
unique(Seurat::Idents(SeuratObj))
```

It's convenient to transfer annotation from reference to query data:

```{r}
df <- predictFun(query_SeuratObj = SeuratObj, reference_SeuratObj = pbmc3k.final,
                 group_by = 'seurat_annotations', cluster_by = 'seurat_clusters',
                 species = "Homo sapiens", by = 'GO', k = NULL, LDAmethod = "VEM")
df
```

To obtain the accuracy of prediction in this case, we identify the cell types of each cluster in query data based on ground-truth identity of each cell.

```{r}
mm <- table(SeuratObj$seurat_annotations, SeuratObj$seurat_clusters)
nn <- setNames(rownames(mm)[apply(mm, 2, which.max)], colnames(mm))
caret::confusionMatrix(data = as.factor(df$prediction), reference = as.factor(nn[df$query]))$overall[["Accuracy"]]
```

To improve accuracy of prediction, users can oversample the reference data. In this case, we skip this step because it can be time-consuming.

```{r, eval = FALSE}
reference_SeuratObj <- oversample_ref(reference_SeuratObj, number_clusters = 5,
                                      group_by = 'cellType', cluster_by = 'seurat_clusters',
                                      species = "Homo sapiens", by = 'GO', k = NULL, method = "VEM")
predictFun(query_SeuratObj, reference_SeuratObj, group_by = 'seurat_annotations', cluster_by = 'seurat_clusters',
           species = "Homo sapiens", by = 'GO', k = NULL, LDAmethod = "VEM")
# if true labels of query data are provided, you can use caret::confusionMatrix to inspect the confusion matrix and accuracy
caret::confusionMatrix(data = prediction , reference = true_label)
caret::confusionMatrix(data = prediction , reference = true_label)$overall[["Accuracy"]]
```

