---
title: "Transfer topic model and visualization"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transfer topic model and visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  message = FALSE,
  # error = FALSE,
  warning = FALSE,
  fig.align = "center"
)
options(width = 80)
```

```{r}
library(CellFunTopic)
library(magrittr)
```

## Reference topic models

CellFunTopic provides built-in reference topic models for users to take advantage of large landscape datasets such as the human cell landscape (HCL) and the mouse cell atlas (MCA).

```{r, eval = FALSE}
?HCL_ldaOut
?MCA_ldaOut
```

CellFunTopic also provides methods for users to make their own reference topic models. 

## Make a reference topic model and transfer

Let us take `pbmc3k.final` as example to show how to make a reference topic model and transfer. We take the original one as the reference data, and re-cluster it as the query data.

Make a reference topic model out of `pbmc3k.final` with the ground-truth cell type.

```{r, eval = FALSE}
data('pbmc3k.final', package = "pbmc3k.SeuratData")
reference_SeuratObj <- readData(data = pbmc3k.final, type = 'Seurat', species = "Homo sapiens")
reference_SeuratObj <- DetectGeneIDtype(reference_SeuratObj)
SeuratObj.markers <- Seurat::FindAllMarkers(reference_SeuratObj, only.pos = TRUE, min.pct = 0.0001, logfc.threshold = 0.0001, return.thresh=0.9)
slot(object = reference_SeuratObj, name = 'misc')[["Allmarkers"]] <- SeuratObj.markers
reference_SeuratObj <- RunGSEA(reference_SeuratObj, by = 'GO')
k <- length(unique(Seurat::Idents(reference_SeuratObj)))
reference_SeuratObj <- runLDA(reference_SeuratObj, by = "GO", k = k, method = "VEM", SEED = 1234, plot = F)
```

```{r, echo = FALSE}
reference_SeuratObj <- readRDS(file = "../vignette_data/pbmc3k_original.rds")
```

```{r}
table(Seurat::Idents(reference_SeuratObj))
```

Re-analyze and re-cluster `pbmc3k.final` as the query data, each cell owns a new cluster identity.

```{r, eval = FALSE}
SeuratObj <- readData(data = pbmc3k.final, type = 'Seurat', species = "Homo sapiens")
SeuratObj <- CalMTpercent(SeuratObj, by = "use_internal_data")
SeuratObj <- QCfun(SeuratObj, plot = F)
SeuratObj <- RunSeurat(SeuratObj, nPCs = 10, resolution = 1, plot = FALSE)
SeuratObj <- RunGSEA(SeuratObj, by = 'GO')
```

```{r, echo = FALSE}
SeuratObj <- readRDS(file = "../vignette_data/pbmc3k_SeuratObj.rds")
```

```{r}
# check the new clustering
table(Seurat::Idents(SeuratObj))
```


Transfer topic model from reference data to query data.

```{r}
ref_ldaOUt <- reference_SeuratObj@misc$ldaOut
```

```{r, eval = FALSE}
SeuratObj <- transfer_LDA(query_SeuratObj = SeuratObj, ref_ldaOUt = ref_ldaOUt, by = "GO")
```

## Visualize transferred model

```{r, fig.height=7, fig.width=7}
dist <- SeuratObj@misc$dist
gammaDF <- reshape2::melt(dist$topics) %>% magrittr::set_colnames(c("document", "topic", "gamma"))
plot_sankey(gammaDF, topn=1)

```

```{r, fig.height=6, fig.width=6}
mm <- dist$topics
colnames(mm) <- paste0('Topic ', colnames(mm))
pheatmap::pheatmap(mm, cluster_rows = T, cluster_cols = T, angle_col = "0", main = "probability between topic and cluster",
                   color = colorRampPalette(c('white', RColorBrewer::brewer.pal(n=9,name="OrRd")))(100),
                   fontsize_row = 8)
```

## Visualize correlation or comparison

CellFunTopic provides a few methods to visualize correlation or comparison between reference data and query data.
For better comparison, we identify the cell types of query data based on ground-truth identity of each cell.

```{r, fig.height=7, fig.width=7}
mm <- table(SeuratObj$seurat_annotations, SeuratObj$seurat_clusters)
nn <- setNames(rownames(mm)[apply(mm, 2, which.max)], colnames(mm))
rownames(dist$topics) <- nn[rownames(dist$topics)]

sankey_comparison(ldaOut=ref_ldaOUt, dist=dist, a_prefix="reference_", b_prefix="query_", height=800)
```

```{r, fig.height=10, fig.width=10}
cosine_network_cluster2(ldaOut=ref_ldaOUt, dist=dist, a_prefix="reference_", b_prefix="query_",
                        layout="fr", cos_sim_thresh=0.5, radius=0.1, text_size = 3)
```

```{r, fig.height=7, fig.width=7}
cosine_heatmap2(ldaOut=ref_ldaOUt, dist=dist, a_prefix="reference_", b_prefix="query_", method="cosine")
```

```{r, fig.height=7, fig.width=7}
cosine_hive(ldaOut=ref_ldaOUt, dist=dist, a_prefix="reference_", b_prefix="query_")
```



